---
title: "Assignment 2- Supplementry"
author: "Paria Fakhrzad"
date: "10/3/2021"
output: 
    pdf_document:
      toc: true
bibliography: Assignment2.bib
fontsize: 11pt
---

# Part1- Data
## 1.1- loading libraries
```{r eval=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(Hmisc)
library(magrittr)
library(readr)
library(ggplot2)
library(ISLR2)
library(class)
library(ggpubr)
library(GGally)
library(PreProcess)
library(caTools)
library(caret)
library(tree) # CART 
library(MASS)
```

## 1.2- loading dataset
The source of data^[https://www.kaggle.com/racholsan/customer-data].
we will use a Car insurance claim dataset in 2021 that has collected based on customers of that insurance company ~\cite{data}. There are 10000 customer as sample in this data frame. Also 17 features have been collected based on the customer's attributes. There is one column that shows this customer had accident claim last year or not. this is logical column. The data shows that 23percent of customers claimed the accident insurance last year.

```{r}
customer_claim_original <- readr:: read_csv("customer-data.csv")
customer_claim <- customer_claim_original
customer_claim <- dplyr::select(customer_claim,-id)
```

## 1.3- Tidying data
The rows with NA values in SpeedingViolation have been removed. Also the age and experience were interval that both have converted to Number(by mean of interval). Other char columns(gender, married,education,...), now are factors with this logic:
* Gender female=1 , male=0 
* married True=1 , false=0 
* children True=1 , false=0

```{r}

#finding null observation
table(sapply(customer_claim,function(x)all(is.na(x))))#Columns are totally empty
table(lapply(customer_claim,function(x){length(which(is.na(x)))})) #Columns with NA
customer_claim<- dplyr::filter(customer_claim , !is.na(annual_mileage))
customer_claim<- dplyr::filter(customer_claim , !is.na(credit_score))

```

```{r}
# changing the age to numbers
customer_claim$age <- ifelse(customer_claim$age=="65+",65,customer_claim$age) 
customer_claim$age <- ifelse(customer_claim$age=="16-25",20,customer_claim$age)
customer_claim$age <- ifelse(customer_claim$age=="26-39",33,customer_claim$age)
customer_claim$age <- ifelse(customer_claim$age=="40-64",52,customer_claim$age) 
customer_claim$age <- as.numeric(customer_claim$age)
```


```{r}

# changing the driving experience to number
customer_claim$driving_experience <- ifelse(customer_claim$driving_experience=="0-9y",5,
                                            customer_claim$driving_experience)
customer_claim$driving_experience <- ifelse(customer_claim$driving_experience=="10-19y",15,
                                            customer_claim$driving_experience)
customer_claim$driving_experience <- ifelse(customer_claim$driving_experience=="20-29y",25,
                                            customer_claim$driving_experience)
customer_claim$driving_experience <- ifelse(customer_claim$driving_experience=="30y+",35,
                                            customer_claim$driving_experience)
customer_claim$driving_experience <- as.numeric(customer_claim$driving_experience)

```


```{r}
#change gender childrean and married to number
customer_claim$gender <- as.numeric(ifelse(customer_claim$gender=="female",1,0))
customer_claim$married <- as.numeric(ifelse(customer_claim$married=="TRUE",1,0))
customer_claim$children <- as.numeric(ifelse(customer_claim$children=="TRUE",1,0))
customer_claim$vehicle_ownership <- as.numeric(
  ifelse(customer_claim$vehicle_ownership=="TRUE",1,0))

```

```{r}
#income
customer_claim$income <- ifelse(customer_claim$income=="upper class",
                                           1,customer_claim$income)
customer_claim$income <- ifelse(customer_claim$income=="middle class",
                                           2,customer_claim$income)
customer_claim$income <- ifelse(customer_claim$income=="working class",
                                           3,customer_claim$income)

customer_claim$income <- ifelse(customer_claim$income=="poverty",
                                           4,customer_claim$income)
customer_claim$income <-as.numeric(customer_claim$income)
#education
customer_claim$education <- ifelse(customer_claim$education=="none",
                                           0,customer_claim$education)
customer_claim$education <- ifelse(customer_claim$education=="high school",
                                           1,customer_claim$education)
customer_claim$education <- ifelse(customer_claim$education=="university",
                                           2,customer_claim$education)
customer_claim$education <-as.numeric(customer_claim$education)
#race
customer_claim$race <- ifelse(customer_claim$race=="minority",0,1)

customer_claim$vehicle_year <- as.factor(customer_claim$vehicle_year)
customer_claim$vehicle_type<- as.factor(customer_claim$vehicle_type)
```

```{r}
#change the label to factor
customer_claim <- mutate(customer_claim, outcome = 
                        factor(outcome, levels = c("FALSE", "TRUE"), 
                        labels = c(0, 1)))
```


# Part2- Data exploration 

## 2.1- Summary of variables
* `customer_claim` is a data frame with 10000 observations on 18 variables.
 * There are eighteen variables in the data set
  * outcome $label$ is a factor column that has two labels, false(has claimed) or true(has not claimed)
  * age $x_{1}$, a range variable that shows the age of customer is in which interval
  * gender  $x_{2}$,
  * race $x_{3}$,
  * driving_experience $x_{4}$,
  * income $x_{5}$,
  * credit_score $x_{6}$,
  * vehicle_ownership $x_{7}$,
  * vehicle_year $x_{8}$,
  * married $x_{9}$,
  * children $x_{10}$,
  * annual_mileage $x_{9}$,
  * vehicle_type $x_{10}$,
  * speeding_violations $x_{11}$,
  * past_accidents $x_{12}$,
```{r}
#shows the feature names
customer_claim <- as_tibble(customer_claim)
names(customer_claim)
dim(customer_claim)
str(customer_claim)

xtable::xtable(summary(dplyr::select(customer_claim,c(age,gender,driving_experience
               ,married,children,postal_code,education))))

```

## 2.2- Analyzing data
In this section we tried to explore the association by visualization tools and some statistical evidence.

### 2.2.1- Using visualization methods
we use ggpairs() \cite{@Emersonetc2012} to see correlation between numeric features in figure1. the significant point is that Age and driving experience are collinear.  In figure4 It appears that older  customers who own the vehicle tended to not claim the insurance for accident(outcome in figure legend) than those who did not. Also in figure5 it seems there are relationship between credit score and number of claims in last year.
* using ggpairs() for founding the correlation of numeric features:
```{r}
customer_corr_matrix <- dplyr::select(customer_claim,c("age","driving_experience","annual_mileage",
                                                       "speeding_violations","past_accidents","credit_score","outcome"))

figure <- ggpairs(customer_corr_matrix)
figure1 <-ggpairs(customer_corr_matrix, columns = 1:7, 
                  ggplot2::aes(colour=outcome))
ggsave("figure1.png",figure1, width=10, height = 4)
figure1
figure

```

* using box-plot and scatter-plot for figuring out the relationships between some categorical variables with "response" that here is ourcome( 1:TRUE, 0:FALSE)
```{r}

# speeding_violations and past accidents
figure2 <-ggscatter(customer_claim,x ='speeding_violations',y ='past_accidents', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson")+
          scale_color_manual(values = c("green", "red"))

# income and age
figure3 <- ggplot(customer_claim) +
  geom_boxplot(aes(x = income, y = age, fill = outcome)) +
  theme(legend.position = "none") +
  theme_bw() +
  scale_fill_manual(values = c("blue2", "orange"))

# vehicle ownership and age
figure4 <- ggplot(customer_claim) +
  geom_boxplot(aes(x = vehicle_ownership, y = age, fill = outcome)) +
  theme(legend.position = "none") +
  theme_bw() +
  scale_fill_manual(values = c("blue2", "orange"))

# credit score and outcome
figure5 <- ggplot(customer_claim) +
  geom_boxplot(aes(x = outcome, y = credit_score)) +
  theme(legend.position = "none") +
  theme_bw()


figure2
figure3
figure4
figure5

ggsave("figure2.png",figure2,width = 3, height = 2)
ggsave("figure3.png",figure3,width = 3, height = 2)
ggsave("figure4.png",figure4,width = 3, height = 2)
ggsave("figure5.png",figure5,width = 3, height = 2)
```

### 2.2.2- Using statistical method
We also have used cor() function to calculated the correlation between features.

```{r}
cor(customer_claim$annual_mileage, customer_claim$speeding_violations, method = "pearson", use = "complete.obs")

```
we used xtabs() to verify that all levels of factor variables had both claimed insurance.
```{r}
# calculatin the number of response per each categorical variable
xtabs(~outcome+age,customer_claim)
xtabs(~outcome+gender,customer_claim)
xtabs(~outcome+race,customer_claim)
xtabs(~outcome+driving_experience,customer_claim)
xtabs(~outcome+education,customer_claim)
xtabs(~outcome+income,customer_claim)
xtabs(~outcome+vehicle_ownership,customer_claim)
xtabs(~outcome+vehicle_type,customer_claim)
xtabs(~outcome+vehicle_year,customer_claim)
xtabs(~outcome+married,customer_claim)
xtabs(~outcome+children,customer_claim)
```

# Part3- Classification
## 3.1- Validation set
*We will use KNN to predict the label of outcome that shows if customer had insurance claim or not. for this mean we need two split our data as train and test.

* train_data: split randomly by sample (75%)
* test_data: split randomly by sample (25%)
```{r}
#defining sample size and randomly split data solution 1
training_size = floor(0.75*nrow(customer_claim))
train_data_Solution1 <- sample(seq_len(nrow(customer_claim)),size = training_size)
test_data_Solution1 <- customer_claim[-train_data_Solution1,]
train_data_Solution1 <- customer_claim[train_data_Solution1,]

#defining sample size and randomly split data solution 2
 ts_split_Solution2 <- createDataPartition(customer_claim$outcome, p = 0.75, list = FALSE)
 train_data_Solution2<- customer_claim[ts_split_Solution2,]
 test_data_Solution2<- customer_claim[-ts_split_Solution2,]

#definign traing and test data solution 3
train_data <- customer_claim %>% 
  mutate(ind=1:nrow(customer_claim)) %>% 
  group_by(outcome) %>% 
  mutate(n=n()) %>%
  sample_frac(size=.75 , weight=n)%>% 
  ungroup()
train_ind<- train_data$ind
test_data <- customer_claim[-train_ind,]


#defining the column label for train and test
label_train_claim <- dplyr::pull(train_data, outcome)
label_test_claim <- dplyr::pull(test_data, outcome)
predictor_train <-dplyr::select(train_data,age,driving_experience,
                annual_mileage,past_accidents,speeding_violations)
predictor_test <- dplyr::select(test_data,age,driving_experience,                         annual_mileage,past_accidents,speeding_violations)

```

## 3.2- KNN
 we fit the model with training data then we predict the response with test data.
 For this classification model we need K, that one way is using $square root of number pf samples$ based on Thumb rule in this link^[https://discuss.analyticsvidhya.com/t/how-to-choose-the-value-of-k-in-knn-algorithm/2606/13]. Also after fitting KNN model We can test our accuracy for K range from 1 till 101.
 * Repeat the K in KNN model to choose the best K
```{r}
accuracy_table <- data.frame(matrix(ncol = 2, nrow= 0))
col_name <- c('accuracy', 'k')
colnames(accuracy_table) <- col_name
for(i in seq(from = 1, to = 101, by = 2)){
  knn_model1 <- knn(predictor_train, predictor_test, train_data$outcome,k = i)
  accuracy <- mean(knn_model1 == label_test_claim)
  accuracy_table[i,'accuracy'] <- accuracy
  accuracy_table[i, 'k'] <- i
}
accuracy_table <- accuracy_table[order(accuracy_table$accuracy, decreasing = TRUE),]
head(accuracy_table,20)
xtable::xtable(head(accuracy_table,20))
```

### 3.2.1 fitting KNN model
```{r}
#fit
knn_model <- knn(predictor_train, predictor_test, train_data$outcome,k = 9)
#predict
KNN.predicted.data <-data.frame(knn_model, label_test_claim)

#Intepret the outptut
summary(knn_model)
```

## 3.2- Logistic regression 
We fit model logistic regression based on the train and test data and run it 10 times. Also we consider that labels with fitted probability more than 50\% are correct and we plot the accuracy probability with samples as figure6. It clearly shows that the logistic regression model works well based on this dataset. Also based on figure8 it shows there is not much outlier.

```{r}
#fit
LR_model<- glm(outcome~driving_experience+
                vehicle_year,
                data =  customer_claim,
                family ="binomial",
                subset = train_ind)

#predict test data 
LR_predict <-predict(LR_model, newdata = test_data, type = "response")

LR.predicted.data <-data.frame(probability=LR_predict, outcome=label_test_claim)
LR.predicted.data <-LR.predicted.data[order(LR.predicted.data$probability, decreasing = FALSE),]
LR.predicted.data$rank <- 1:nrow(LR.predicted.data)
cutoff <- 0.5
LR.predicted.data <- dplyr::mutate(LR.predicted.data, predicted_outcome = ifelse(LR.predicted.data$probability > cutoff,1,0)) 

#Intepret the outptut
summary(LR_model)
summary(LR_model)$r.sq
ggsave ("figure8.png",plot(LR_model))
#xtable::xtable(vif(LR_model))
```


* here we plot the probability per response for Logistic regression model
```{r}
# Probability per response 
figure7 <- ggplot(LR.predicted.data, aes(x= rank, y=probability, col=outcome))+
           geom_point()+scale_fill_manual(values = c("red", "green"))
figure7
ggsave("figure7.png",figure7,width = 3, height = 2)
```

## 3.3- Classification Decision tree 
### 3.3.1  Fit classification tree
```{r}
Tree_model <- tree(outcome~age+gender+education+married+income+
                  driving_experience+annual_mileage+vehicle_year+
                  vehicle_type+speeding_violations+past_accidents+
                  children,
                  data=train_data)
Tree_predict <- predict(Tree_model, test_data, type = "class")
set.seed(7)

#Pruning
Tree_cv <- cv.tree(
  Tree_model, 
  FUN = prune.misclass)
Tree_cv
plot(Tree_cv$size, Tree_cv$dev, type = "b") 
plot(Tree_cv$k, Tree_cv$dev, type = "b")

prune.Tree <- prune.misclass(Tree_model, best = 3)  

plot(prune.Tree)
text(prune.Tree, pretty = 0)
Tree_predict <- predict(prune.Tree, test_data, type = "class")

#Intepret the outptut
summary(Tree_model)
plot(Tree_model)
text(Tree_model, pretty = 0)
```
## 3.4- Model Evaluation

### 3.4.1- Confusion matrix
```{r}
# KNN accuracy
table(knn_model, label_test_claim)
mean(knn_model == label_test_claim)

#Logistic regression accuracy
table(LR.predicted.data$predicted_outcome,label_test_claim)
mean(LR.predicted.data$predicted_outcome == LR.predicted.data$outcome)

#classification decision tree
table(Tree_predict, label_test_claim)
(table(Tree_predict, label_test_claim)[1,1] + table(Tree_predict, label_test_claim)[2,2])/sum(table(Tree_predict, label_test_claim))

```
that shows accuracy of Logistic regression is a litter more. The total accuracy is around 82\%  which shows that there is more opportunity to improve the model performance.

## 3.5- Cross Validation: validation set
repeat KNN and logistic regression models for 10 times. here we used for() function for iteration.
```{r}
KNN_running_table <- data.frame(matrix(ncol = 2, nrow= 0))
LR_running_table <- data.frame(matrix(ncol = 2, nrow= 0))
colnames(KNN_running_table) <- c('Accuracy', 'Run')
colnames(LR_running_table) <- c('Accuracy', 'Run')
for(i in seq(from = 1, to = 10, by = 1)){
  
  #Splitting data
  train_data <- customer_claim %>% 
  mutate(ind=1:nrow(customer_claim)) %>% 
  group_by(outcome) %>% 
  mutate(n=n()) %>%
  sample_frac(size=.75 , weight=n)%>% 
  ungroup()
  train_ind2<- train_data$ind
  test_data <- customer_claim[-train_ind2,]
  
  predictor_train <-dplyr::select(train_data,age,driving_experience,
                               annual_mileage,past_accidents,speeding_violations)
  predictor_test <- dplyr::select(test_data, age, driving_experience,
                               annual_mileage,past_accidents,speeding_violations)
  label_train_claim <- dplyr::pull(train_data, outcome)
  label_test_claim <- dplyr::pull(test_data, outcome)
  
  #knn model
  knn_model2<- knn(predictor_train, predictor_test, train_data$outcome,k = 43)
  
  #Logistic regression model
  LR_model <- glm(outcome ~age+education+income+postal_code+driving_experience+
                annual_mileage+vehicle_year+vehicle_type+
                speeding_violations+past_accidents,
                data=customer_claim,
                family = binomial("logit"),
                subset =train_ind)
  LR_predict <-predict(LR_model, newdata = test_data, type = "response")
  LR_predicted_table <-ifelse(LR_predict >= 0.5,1,0)
  
  #filling the accuracy table
  KNN_accuracy <- mean(knn_model2 == label_test_claim)
  KNN_running_table[i,'Accuracy'] <- KNN_accuracy
  KNN_running_table[i, 'Run'] <- i
  LR_accuracy <- mean(LR_predicted_table == label_test_claim)
  LR_running_table[i,'Accuracy'] <- LR_accuracy
  LR_running_table[i, 'Run'] <- i
}
#making table from output and calculating the average
xtable::xtable( KNN_running_table)
mean(KNN_running_table$Accuracy)
xtable::xtable( LR_running_table)
mean(LR_running_table$Accuracy)

```
## 3.6- Cross Validation: K-Fold
* K-fold CV for Logistic regression
```{r}

library(boot)
LR_cv<- glm(outcome ~age+income+driving_experience+annual_mileage+
                vehicle_year+vehicle_type+speeding_violations+past_accidents, data = customer_claim, family = "binomial")
1-cv.glm(customer_claim, LR_cv, K=29)$delta[1]
```
* K-fold CV for KNN
```{r}
library(class)
library(e1071)
x <- dplyr::select(customer_claim,age,driving_experience,
            annual_mileage,past_accidents,speeding_violations)
y <- dplyr::pull(customer_claim,outcome)
knn_cv <- tune.knn(x,y, k = 1:30, tunecontrol = tune.control(sampling = "cross",cross=5))
summary(knn_cv)
plot(knn_cv)
```

# Part4- Clustering
## 4.1- PCA
```{r}
#remove categorical features 
customer_claim_num <- dplyr::select(customer_claim,-vehicle_year)
customer_claim_num <- dplyr::select(customer_claim_num,-vehicle_type)
customer_claim_num <- dplyr::select(customer_claim_num,-outcome)

#check the data
str(customer_claim_num)

#perform PCA
PCR_model <-prcomp(customer_claim_num,center= TRUE, scale = TRUE)
names(PCR_model)
PCR_model$rotation
dim(PCR_model$x)
summary(PCR_model)
plot(PCR_model$x)
plot(PCR_model)
```

```{r}
biplot(PCR_model, scale=0)
```
```{r}
PCR_model$rotation=-PCR_model$rotation
PCR_model$x=-PCR_model$x
biplot(PCR_model, scale=0)
```
```{r}
prop_var <- (PCR_model$sdev^2)/sum(PCR_model$sdev^2)
prop_var
plot(prop_var, type="b")
plot(cumsum(prop_var), type="l")
```


## 4.2- Kmeans clustering 
```{r}
KM_model <-kmeans(customer_claim_num, 3, nstart=2)
KM_model
KM_model$tot.withinss
```

```{r}

```


## 4.3- Hierarchical Clustering

```{r}
customer_df_dis<- dist(customer_claim_num,method = 'euclidean')
hcl_com_model <-hclust(customer_df_dis, method = "complete")
hcl_ave_model <-hclust(customer_df_dis, method = "average")
hcl_sin_model <-hclust(customer_df_dis, method = "single")
hcl_cen_model <-hclust(customer_df_dis, method = "centroid")

```

```{r}
plot(hcl_com_model, main = "Complete Linkage",
xlab = "", sub = "", cex = .5, hang=-10)
rect.hclust(hcl_com_model , k = 3, border = 2:6)
abline(h = 3, col = 'red')

plot(hcl_ave_model, main = "Average Linkage",
xlab = "", sub = "", cex = .5, hang=-10)
rect.hclust(hcl_ave_model , k = 3, border = 2:6)
abline(h = 3, col = 'red')

plot(hcl_sin_model, main = "Single Linkage",
xlab = "", sub = "", cex = .5, hang=-10)
rect.hclust(hcl_sin_model , k = 3, border = 2:6)
abline(h = 3, col = 'red')

plot(hcl_cen_model, main = "centroid Linkage",
xlab = "", sub = "", cex = .5, hang=-10)
rect.hclust(hcl_cen_model , k = 3, border = 2:6)
abline(h = 3, col = 'red')
```
```{r}
cut_hclc_model<-cutree(hcl_com_model,3)
cut_hcla_model<-cutree(hcl_ave_model,3)
cut_hcls_model<-cutree(hcl_sin_model,3)
cut_hclcen_model<-cutree(hcl_cen_model,3)
```

```{r}
customer_hcl<- mutate(customer_claim_num,cluster=cut_hclc_model)
ggplot(customer_hcl, aes(x=customer_hcl$annual_mileage, y =customer_hcl$credit_score, color = factor(cluster))) + geom_point()

```

```{r}
customer_scale<- scale(customer_claim_num)
plot(
  hclust(dist(customer_scale), method = "complete"),
main = "Hierarchical Clustering with Scaled Features"
)
```
## 4.4- DBSCAN clustering 
In this algorithm there is no need to specify the K, 

## 4.5- MeanShift clustering


## 4.6- Spectral Clustering


## 4.7- Gaussian Mixture Model clustering
