---
title: "Assignment1-Data Visualization and tidy"
author: "Paria Fakhrzad"
date: "9/18/2021"
output: 
  pdf_document:
    toc: true
bibliography: Assignment1.bib
fontsize: 12pt
---
\newpage
# loading libraries
in this assignment I use of this dataset  noaa_gsod^[https://gist.githubusercontent.com/krisrs1128/3845514e2d5eef57ec3271ea20fdcdb1/raw/44f9860df64b280e6b4c155f144c67bea1825edd/noaa-gosd.csv]
for more information about this data set fo through this link^[https://www.kaggle.com/noaa/gsod].

 * These libraries have been used
```{r eval=FALSE, warning=FALSE , message=FALSE}
library(dplyr)     # using for adding columns or setting filter
library(readr)     # reading data
library(tidyverse) 
library(tidyr)    # using gather for tidying
library(ggplot2)  # plots
library(pheatmap) # plot heatmap
library(readxl)   # save file as CSV
library(phyloseq)
library(lubridate)# converting the string to dates and extract year and month
library(viridis)  # put color in plots
library(shiny)    # building shiny app
library(xtable)   # making table for LATEX
library(plyr)     # use for calculating frequency
library(shinyWidgets)
```
# Assignment1
## Part a
I am using the ggplot2 library for more information you can go through this [link](https://ggplot2.tidyverse.org/)

 * the data has loaded and the geom_hex is used for making a histogram
```{r eval=FALSE}
data_weather <- readr:: read_csv("https://gist.githubusercontent.com/krisrs1128/3845514e2d5eef57ec3271ea20fdcdb1/raw/44f9860df64b280e6b4c155f144c67bea1825edd/noaa-gosd.csv")
plot1a <- ggplot(data = data_weather) +
    geom_hex(mapping =aes(x=gust, y=visib),binwidth=20)
figure1a <- ggsave("figure1a.png",plot=plot1a, width=10, height=3)
  
```

## Part b
in this part I use filter function to make a new object with clean data in visible and gust columns.
```{r eval=FALSE}
data_weather_2 <- dplyr :: filter(data_weather, visib!=999.9)
data_weather_2 <- dplyr :: filter(data_weather_2, gust!=999.9)
```

## Part c
in this part I create the histogram again with cleaned data to see what would change
```{r eval=FALSE}
figure1b <-ggplot(data = data_weather_2) +
  geom_hex(mapping =aes(x=gust, y=visib),binwidth=1)
ggsave("figure1b.png",plot=figure1b, width=10, height=3)
```

\newpage
# Assignment2

## Part a
in this part I load the data and create a heatmap from raw data with pheatmap()[@pheatmap]
```{r}
data_gene <- readr::read_csv("https://gist.githubusercontent.com/krisrs1128/b8dc85b659186259715f8efe950ffce6/raw/77669ebdda4d6f8a029c2fdf506f4599277b50f4/hiiragi.csv")

# finding the null data 
is.na(data_gene) %>% table()
dim(data_gene)

#try to remove the first column as gene and put it as row names
data_gene2 <- data_gene[,1]
df <-dplyr::select(data_gene,-gene)
data_gene_tied <- dplyr::select(data_gene,-gene)
df2 <- as.matrix.data.frame(data_gene)
rownames(df2) <- data_gene2$gene
df2 <- dplyr::select( data_gene, "1 E3.25" : "101 E4.5 (FGF4-KO)")


pheatmap1<- pheatmap(df2,cluster_rows=FALSE,cluster_cols=FALSE)
ggsave("figure2a.png",plot=pheatmap1,width=10, height=4)

```


## Part b
in this part I use kmeans_k as number of clusters and also put denogram for both rows and columns. I see one category in samples names but I am not sure it works or not.

```{r}
rownames(df2) <- data_gene2$gene
pheatmap3<- pheatmap(df2,cluster_rows=TRUE, cluster_cols=TRUE, kmeans_k = 10,)
ggsave("figure2b.png",plot=pheatmap3,width=10, height=4)
pheatmap3
```
\newpage
# Assignment3

## Part a and b

 * reading data
in this Assignment I used the data in this [Link](https://www.kaggle.com/shivkp/customer-behaviour?select=DataCoSupplyChainDataset.csv). this is a table with 180519 rows and 53 columns. 

```{r}
SCM_data <- read_csv("DataCoSupplyChainDataset.csv")
```

* Tidying the Supply Chain data
note: In this table, one column has a NULL number. I found this based on below codes:
```{r}
summary_SCM <-dplyr::select(SCM_data,c("Days for shipping (real)","Sales per customer","Sales per customer","Order Item Discount Rate","Type"))
xtable::xtable(summary_SCM <- data.frame(summary(summary_SCM))) #table of summaryS
sapply(SCM_data,function(x)all(is.na(x))) #Columns are totally empty
table(sapply(SCM_data,function(x)all(is.na(x))))#Columns are totally empty
table(lapply(SCM_data,function(x){length(which(is.na(x)))})) #Columns with NA
str(SCM_data)

```

* spliting tables
because we don't want to use all columns so some of them has been selected and the names of columns have changed based on standard valuable names in programming( removing the spaces between words)

```{r}
#changing the name of columns
names(SCM_data)[names(SCM_data) == "Delivery Status"] <- "delivery_status"
names(SCM_data)[names(SCM_data) == "Shipping Mode"] <- "shipping_mode"
names(SCM_data)[names(SCM_data) == "Days for shipping (real)"] <- "shipping_days"
names(SCM_data)[names(SCM_data) == "order date (DateOrders)"] <- "order_date"
names(SCM_data)[names(SCM_data) == "shipping date (DateOrders)"] <- "shipping_date"
names(SCM_data)[names(SCM_data) == "Category Name"] <- "category_name"
names(SCM_data)[names(SCM_data) == "Order Item Quantity"] <- "order_quantity"
names(SCM_data)[names(SCM_data) == "Order Item Discount"] <- "discount"
names(SCM_data)[names(SCM_data) == "Order Item Discount Rate"] <- "discount_rate"
names(SCM_data)[names(SCM_data) == "Order Item Product Price"] <- "price"
names(SCM_data)[names(SCM_data) == "Order Region"] <- "order_region"
names(SCM_data)[names(SCM_data) == "Benefit per order"] <- "benefit_order"
names(SCM_data)[names(SCM_data) == "Order Profit Per Order"] <- "Profit"
names(SCM_data)[names(SCM_data) == "Department Name"] <- "department"

#changing date
SCM_data <- mutate(SCM_data, 
                      order_date = as.Date(order_date, format="%m/%d/%Y"),
                      shipping_date = as.Date(shipping_date, format="%m/%d/%Y"),
                      year = year(order_date),
                      month=month(order_date)
                      )
#selecting the columns for ploting
table_delay <-  dplyr::select(SCM_data,c("delivery_status","shipping_mode","shipping_days","order_date","shipping_date","category_name","Type","department","year","month"))

```


## Part c
 * graph number 1
```{r}
sub_group <-dplyr::filter(table_delay,year==c(2015,2016,2017,2018))
graph_bar <- ggplot(sub_group)+
  geom_bar(aes(order_date,fill = factor(sub_group$delivery_status)))

ggsave("figure3a.png",plot=graph_bar, width=10, height=4)

```


```{r}
table_price <- dplyr::select(SCM_data,c("Late_delivery_risk","year","shipping_date","shipping_days","discount","discount_rate","price","order_quantity","order_region","shipping_mode",))

graph_point <- ggplot(table_price)+
    geom_point(
    aes(x= shipping_date,y= shipping_days, 
    colour= factor(Late_delivery_risk)))+
    scale_color_viridis(discrete = TRUE) +
    facet_wrap(.~table_price$order_region)
 ggsave("figure3b.png",plot=graph_point, width=10, height=4)
```

```{r}
graph_his <- ggplot(data = SCM_data) +
  geom_hex(mapping =aes(x=Profit, y=discount),binwidth=20)
ggsave("figure4c.png",plot=graph_his, width=10, height=4)

```

* Graph number3
 comparing shipment class and Delivery status. I split new table that contains two variables and add new column for frequency.
```{r}
matrix_data <- select(table_delay,shipping_mode,delivery_status)
counts <- ddply(matrix_data, .(matrix_data$shipping_mode, matrix_data$delivery_status), nrow)
names(counts)[names(counts) == "matrix_data$shipping_mode"] <- "shipping_class"
names(counts)[names(counts) == "matrix_data$delivery_status"] <- "shipping_status"
names(counts)[names(counts) == "V1"] <- "frequency"

graph_bar <- ggplot(data=matrix_data, aes(x=shipping_mode,, fill=factor(delivery_status)))+
                  geom_bar()
ggsave("figure3c.png",plot=graph_bar, width=10, height=4)
```


## Part d- Shiny App
In this Part I used the customize dataset in my github repository in this [link](https://raw.githubusercontent.com/pfakhrzad/Data_visualization/main/SCM_shiny.csv)

```{r}

write.csv(sub_group,'SCM_shiny.csv')
data_for_shiny <- read_csv("https://raw.githubusercontent.com/pfakhrzad/Data_visualization/main/SCM_shiny.csv")
library(shiny)
library(shinyWidgets)

```

* making the shiny app 
the link of shiny app is[https://vmbe7g-paria-fakhrzad.shinyapps.io/Smart_Supply_chain/](https://vmbe7g-paria-fakhrzad.shinyapps.io/Smart_Supply_chain/)

```{r}
# Define UI part 
ui <- fluidPage(
  setBackgroundColor("gray"),
  
# Application title
        titlePanel("Delayed_shipments"),

# A select Input for product category
        selectInput(inputId ="ID",
                    label   ="Choose the Product category",
                    choices =c("Fitness","Apparel","Golf","Footwear","Outdoors",
                               "Fan Shop","Technology","Book Shop","Discs Shop"
                                ,"Pet Shop","Health and Beauty"),
                    selected="Apparel",width=200
                   ),
        
# Show a plot of the Delayed shipment
        plotOutput("bargraph")
               )

# Define server logic required to draw a bargragh
server <- function(input, output)
{

    output$bargraph <- renderPlot({
      library(dplyr)
      SelectedData <- dplyr::filter(data_for_shiny,department==input$"ID")
      library(ggplot2)
      title <- "Delayed_Shipment"
      ggplot(SelectedData)+
          geom_bar(aes(order_date,colour = factor(delivery_status)))+
          ggtitle("The number of orders trend with delivery status")
                                })
}  

# Run the application 
shinyApp(ui = ui, server = server)
```



